---
output:
  pdf_document: default
  html_document: default
---

# Are Danes having fun in Denmark?

Semester project for Statistics course.

Author: Adam Wolkowycki

```{r}
library(plyr)
# install.packages("GGally")
library(tidyverse)
library(GGally)
library(ggthemes)
options(repr.plot.width=32, repr.plot.height=20)
```
```{r}
read_csv("./archive/world-happiness-report-2021.csv") -> data_2021_raw
read_csv("./archive/world-happiness-report.csv") -> data_all_raw

head(data_2021_raw)
head(data_all_raw)
```

Combine two dataframes

```{r}
# Changing the column names of data_all_raw and naming the dataframe data_all
data_all_raw %>% select(country = 'Country name', score = 'Life Ladder', 
                        economy = 'Log GDP per capita',
                        social_support = 'Social support',
                        life_expectancy = 'Healthy life expectancy at birth',
                        freedom = 'Freedom to make life choices',
                        generosity = 'Generosity',
                        corruption = 'Perceptions of corruption', year) -> data_all

# Changing the column names of data_2021_raw and naming the dataframe data_all
data_2021_raw %>% select(country = 'Country name', score = 'Ladder score', 
                        economy = 'Logged GDP per capita',
                        social_support = 'Social support',
                        life_expectancy = 'Healthy life expectancy',
                        freedom = 'Freedom to make life choices',
                        generosity = 'Generosity',
                        corruption = 'Perceptions of corruption',
                        region = 'Regional indicator') -> data_2021

data_2021 %>% select(country, region) -> continent

full_join(data_all, continent, by = "country") -> data_all

bind_rows(data_all, data_2021) -> data

data[,"year"][is.na(data[,"year"])] <- 2021

head(data)
```

## Denmark

```{r}
dk <- data %>% filter(country == "Denmark")

ggplot(dk, aes(x = year, y = score, label = score)) + geom_point(size = 8, color = "black") +
  theme_fivethirtyeight() +
  geom_segment(aes(x = year, xend = year, y = 0, yend = score)) + 
  scale_x_continuous(breaks = seq(2005, 2021, 1)) +
  scale_y_continuous(breaks = seq(0, 7, 1)) + geom_text(color = 'white', size = 2) +
  labs(title = "Are Danes having fun in Denmark?",
       subtitle = "Happiness scores of Danes across the years",
       x = "Year", y = "Happiness score") +
  theme(plot.title = element_text(size = 15, face = "bold"),
        plot.subtitle = element_text(size = 10),
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 8)) +
  coord_flip()
```

After seeing the scores, I want to see the linear relationship between score and other variables.

```{r}
dk_corr <- dk %>% select(-c(country, year, region))

ggcorr(dk_corr,
       method = c("everything", "pearson"),
       size = 3, hjust = 0.77,
       low = "#ff0000", mid = "white", high = "#2e8b57",
       label = TRUE, label_size = 3,
       layout.exp = 1) + theme_fivethirtyeight()
```

Note: live expectancy and economy have a strong linear correlation.

## Denmark vs Europe

```{r}
happiest_countries_2021 <- data_2021 %>%
    select(country, score) %>%
    group_by(score) %>%
    arrange(desc(score)) %>%
    head(10)

ggplot(happiest_countries_2021, aes(x = country, y = score, label = score)) + 
  geom_point(size = 8, color = "black") +
  theme_fivethirtyeight() +
  geom_segment(aes(x = country, xend = country, y = 0, yend = score)) + 
  geom_text(color = 'white', size = 2) +
  labs(title = "TOP 10",
       subtitle = "Scores of the happpest countries in 2021",
       x = "Year", y = "Happiness score") +
  theme(plot.title = element_text(size = 15, face = "bold"),
        plot.subtitle = element_text(size = 10),
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 8)) +
  coord_flip()
```

## Denmark vs Scandinavia 

```{r}
scandinavia <- data %>% filter(country == 'Denmark' | country == 'Norway' | 
                                 country == 'Sweden' | country == 'Finland' | 
                                 country == 'Iceland') %>%
  group_by(year) %>% 
  mutate(yrrank = row_number(-score))
  
ggplot(scandinavia, aes(x = year, y = score, label = yrrank)) + 
  geom_point(aes(color = country), size = 5) +
  theme_fivethirtyeight() +
  geom_line(aes(color = country)) + scale_x_continuous(breaks = seq(2006, 2021, 1)) +
  scale_y_continuous(breaks = seq(0, 5, 0.5)) + geom_text(color = 'white', size = 3) +
  scale_color_manual(values = c("#b8860b", "#2e8b57", "#66ff00", "#560319", "#efcc00")) +
  labs(title = 'Denmark vs Scandinavia', 
       subtitle = 'Ranking the happiness score of European countries across the years',
       x = 'Year', y = 'Happiness score') + 
  theme(plot.title = element_text(size = 15, face = 'bold'),
        plot.subtitle = element_text(size = 8),
        axis.title = element_text(size = 8),
        axis.text = element_text(size = 8),
        legend.direction = 'horizontal', legend.position = 'bottom',
        legend.title = element_blank(),
        legend.text = element_text(size = 6))
```

## Denmark vs World

```{r}
# Changing the values of region column in Denmark for grasping purposes
data %>% 
  mutate(region = case_when(country == 'Denmark' ~ 'Denmark', TRUE ~ region)) -> data

# Checking the unique and missing values
unique(data['region'])
colSums(is.na(data))

# There are 63 missing values in the column region, so
# investigate null values in this column
data %>% filter(is.na(region)) -> na_region

# Fill up values of the countries with NA region
data %>%
  mutate(
    region = case_when(country =="Angola" | country == "Central African Republic" |
                         country == "Congo (Kinshasa)" | country == "Djibouti"| 
                         country == "Somalia" | country == "Somaliland region"|
                         country == "South Sudan" | country == "Sudan"
                       ~ "Sub-Saharan Africa", TRUE ~ region)) %>% 
  mutate(
    region = case_when(country == "Belize" | country == "Cuba" | country == "Guyana" |
                         country == "Suriname" | country == "Trinidad and Tobago"
                       ~ "Latin America and Caribbean", TRUE ~ region)) %>% 
  mutate(
    region = case_when(country == "Oman" | country == "Qatar" | country == "Syria" 
                       ~ "Middle East and North Africa", TRUE ~ region)) %>% 
  mutate(
    region = case_when(country == "Bhutan" ~ "South Asia", TRUE ~ region)) -> data

# Checking if any country is missed
data %>% filter(is.na(region)) -> na_region

# Combining some regions
revalue(data$region, c("Central and Eastern Europe" = "Europe and North Asia")) -> 
  data$region
revalue(data$region, c("Commonwealth of Independent States" = "Europe and North Asia")) -> 
  data$region
revalue(data$region, c("East Asia" = "West, East, Southeast Asia and North Africa")) -> 
  data$region
revalue(data$region, c("Southeast Asia" = "West, East, Southeast Asia and North Africa")) -> 
  data$region
revalue(data$region, c("Middle East and North Africa" = "West, East, Southeast Asia and North Africa")) -> 
  data$region

# Checking the new set of unique values
unique(data["region"])
```

Creating a new dataframe with average happiness score of each country 
based on their region and also ranking them by year.

```{r}
data %>% 
  select(year, region, score) %>% 
  group_by(year,region) %>% 
  summarise(avg_score = mean(score)) %>%
  mutate_at(vars(region), factor) %>% 
  group_by(year) %>% 
  mutate(yrrank = row_number(-avg_score)) -> avg_region

head(avg_region)
```
```{r}
worldmap = map_data('world') %>% filter(region != 'Antarctica')

merged_data <- merge(x = worldmap, y = data_2021, 
                     by.x = 'region', by.y = 'country', all.x = TRUE) %>% 
  arrange(order)
ggplot(merged_data, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = score)) +

theme_fivethirtyeight() +
  scale_fill_continuous(low = 'white', high = '#2e8b57') +
  labs(title = "World's Average Happiness",
       subtitle = 'Average happiness score of each region in the world in 2021') +
  theme(legend.direction = 'vertical', legend.position = 'right',
        legend.text = element_text(size = 10))
```
```{r}
ggplot(avg_region, aes(x = year, y = avg_score, label = yrrank)) + 
  theme_fivethirtyeight() +
  geom_point(aes(color = region), size = 5) +
  geom_line(aes(color = region)) + scale_x_continuous(breaks = seq(2006, 2021, 1)) +
  scale_y_continuous(breaks = seq(0, 8, 0.5)) + geom_text(color = 'white', size = 3) +
  scale_color_manual(values = c("#be4f62","#ff0000","#191970","#b8860b", 
                                "#2e8b57", "#8a2be2", "#000000", "#66ff00")) +
  labs(title = 'Denmark vs World',
       subtitle = 'Average happiness score of each region in the world across the years',
       x = 'Year', y = 'Happiness score') + 
  theme(plot.title = element_text(size = 15, face = 'bold'),
        axis.title = element_text(size = 8),
        axis.text = element_text(size = 8),
        legend.direction = 'horizontal', legend.position = 'bottom',
        legend.title = element_blank(),
        legend.text = element_text(size = 6))
```

## Denmark vs World in 2021

Now, I am using 2021 data and it doesn't have null values.

```{r}
data %>% filter(year == 2021) -> data_2021_final

colSums(is.na(data_2021_final))

head(data_2021_final)
```

Once again, I want to see the linear relationship between the variables, 
so I am making a correlation matrix.

```{r}
data_2021_final %>% select(-c(country, year, region)) -> data_2021_corr

ggcorr(data_2021_corr,
       method = c("everything", "pearson"),
       size = 3, hjust = 0.77,
       low = "#ff0000", mid = "white", high = "#2e8b57",
       label = TRUE, label_size = 3,
       layout.exp = 1) + theme_fivethirtyeight()
```

Note: Economy, social support and life expectancy are highly correlated to score.

## Inference

```{r}
ggplot(data_2021) +
  geom_point(aes(x = economy, y = score, color = region))
```
```{r}
ggplot(data_2021) +
  geom_histogram(aes(x = economy, fill = region))
```
```{r}
ggplot(data_2021) +
  geom_density(aes(x = score, fill = region), alpha = 0.3)
```
```{r}
xframe <- seq(-10, 10, length = 100)
dnorm(xframe)

sd_world <- sd(data_2021$score)
mean_world <- mean(data_2021$score)
ggplot(data_2021) +
  geom_histogram(aes(x = score, y = ..density.., fill = 'red')) +
  stat_function(fun = dnorm, args = list(mean = mean_world, sd = sd_world))
```
```{r}
ggplot(data_2021) +
  stat_ecdf(aes(x = score), color = 'red', alpha = 0.3) +
  geom_line(stat = 'function', fun = pnorm, args = list(mean = mean, sd = sd))
```
```{r}
ggplot(data_2021) +
  stat_qq(aes(sample = score), color = 'red', alpha = 0.3) +
  stat_qq_line(aes(sample = score))
```

Let's assume that the mean happiness is equal 7.

```{r}
x_dash <- 7
z_score <- (x_dash - mean_world) / sd_world

ggplot(data.frame(x = seq(0, 10, length = 500)), aes(x = x)) +
  stat_function(fun = dnorm, args = list(mean = mean_world, sd = sd_world)) +
  geom_segment(aes(x = x_dash, y = 0, xend = x_dash, 
                   yend = dnorm(x_dash, mean = mean_world, sd = sd_world), color = 'red')) +
  geom_area(stat = 'function', fun = dnorm, args = list(mean = mean_world, sd = sd_world), 
            fill = 'blue', xlim = c(0, x_dash), alpha = 0.5)
```
```{r}
pnorm(z_score)
pnorm(x_dash, mean_world, sd_world)
```

Note: Speaking of the population from a sample, we can say that 
the mean happiness score is less than 7 with 91.4% confidence.

## Linear regression

```{r}
# Merging the Europe to make trend more linear
data_2021 %>%
  mutate(region = case_when(region == 'Central and Eastern Europe' | 
                              region == 'Western Europe' ~ 'Europe', 
                            TRUE ~ region)) -> data_2021_merged

ggplot(data_2021_merged, aes(x = life_expectancy, y = score)) +
  geom_point(aes(color = region), size = 3, alpha = 0.8) +
  geom_smooth(aes(color = region, fill = region), method = 'lm', fullrange = TRUE) +
  facet_wrap(~region) +
  theme_fivethirtyeight() +
  theme(legend.direction = 'horizontal', legend.position = 'bottom',
        legend.title = element_blank(),
        legend.text = element_text(size = 6))
```

## Predictions

Predict the happiness score of 90 year old people from various regions.

```{r}
data_2021_merged %>% filter(region == 'Europe') %>%
  lm(score ~ life_expectancy, data = .) -> eur_fit

data_2021_merged %>% filter(region == 'Latin America and Caribbean') %>%
  lm(score ~ life_expectancy, data = .) -> la_fit

data_2021_merged %>% filter(region == 'North America and ANZ') %>%
  lm(score ~ life_expectancy, data = .) -> na_fit

data_2021_merged %>% filter(region == 'South Asia') %>%
  lm(score ~ life_expectancy, data = .) -> asia_fit

# Europe
summary(eur_fit)$coefficients[1] + summary(eur_fit)$coefficients[2] * 90
# Latin America
summary(la_fit)$coefficients[1] + summary(la_fit)$coefficients[2] * 90
# North America
summary(na_fit)$coefficients[1] + summary(na_fit)$coefficients[2] * 90
# Asia
summary(asia_fit)$coefficients[1] + summary(asia_fit)$coefficients[2] * 90
```
