```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Chi^2 distribution

```{r}
ggplot(data.frame(x = seq(0, 30, length = 100)), aes(x = x)) +
  stat_function(fun = dchisq, args = list(df = 2), color = 'blue') +
  stat_function(fun = dchisq, args = list(df = 3), color = 'green') +
  stat_function(fun = dchisq, args = list(df = 5), color = 'yellow') +
  stat_function(fun = dchisq, args = list(df = 7), color = 'orange') +
  stat_function(fun = dchisq, args = list(df = 11), color = 'red') +
  stat_function(fun = dchisq, args = list(df = 19), color = 'purple') 
```

# Chi^2 goodness of fit test

Sample of 669 Haribo Gummy Bears

```{r}
(gummy_bears <- c(83, 142, 100, 103, 104, 137))
```

H0: Equal proportion of all flavours of gummy bears are produced
H1: Unequal proportion of all flavours of gummy bears are produced

```{r}
alpha <- 0.05 
(expected <- 669 / 6)
```

$$Z_n = \frac{observed\:count - null\:count}{\sqrt{null\:count}}$$

$$X^2 = Z_1^2 + Z_2^2 + \cdots + Z_n^2$$

```{r}
(gummy_Z <- (gummy_bears - expected) / sqrt(expected))

(chi_score <- sum(gummy_Z ^ 2))
```

df = 5

```{r}
ggplot(data.frame(x = seq(0, 30, length = 100)), aes(x = x)) +
  stat_function(fun = dchisq, args = list(df = 5)) +
  geom_vline(xintercept = chi_score, color = 'green')
```

```{r}
(p_value <- 1 - pchisq(chi_score, df = 5))

p_value > alpha
```

We reject hypothesis H0 in favor of the alternative. There is no equal 
proportion in between flavours of gummy bears being produced.

```{r}
chisq.test(x = gummy_bears, p = rep(c(expected), length(gummy_bears)),
           rescale.p = TRUE)
```

# Chi^2 test for independence

```{r}
(movies_pizza <- tribble(
  ~Movie, ~Pepperoni, ~Mushrooms, ~Kebab,
  'Frozen 2', 20, 10, 5,
  'Joker', 15, 12, 15,
  'Someone great', 8, 13, 2
))
```

H0: There is no correlation in between pizza topping and movie
H1: There is correlation in between pizza topping and movie

$$Expected\:count _{\text{row } i, \text{col } j} = \frac{(\text{row } i \text{ total}) \times(\text{column } j \text{ total})}{\text {table total}}$$

```{r}
P_total <- sum(movies_pizza$Pepperoni)
M_total <- sum(movies_pizza$Mushrooms)
K_total <- sum(movies_pizza$Kebab)
(total <- P_total + M_total + K_total)
```
```{r}
movies_pizza <- movies_pizza %>%
  mutate(P_exp = (Pepperoni + Mushrooms + Kebab) * P_total / total) %>%
  mutate(M_exp = (Pepperoni + Mushrooms + Kebab) * M_total / total) %>%
  mutate(K_exp = (Pepperoni + Mushrooms + Kebab) * K_total / total)
```

$$Z_n = \frac{observed\:count - null\:count}{\sqrt{null\:count}}$$
```{r}
movies_pizza <- movies_pizza %>%
  mutate(P_Z = ((Pepperoni - P_exp) / sqrt(P_exp)) ^ 2) %>%
  mutate(M_Z = ((Mushrooms - M_exp) / sqrt(M_exp)) ^ 2) %>%
  mutate(K_Z = ((Kebab - K_exp) / sqrt(K_exp)) ^ 2)

(chi_score_mp <- sum(movies_pizza$P_Z) + sum(movies_pizza$M_Z) + sum(movies_pizza$K_Z))
```

$$d f = (R-1) \times(C-1)$$

df = 4

```{r}
ggplot(data.frame(x = seq(0, 15, length = 100)), aes(x = x)) +
  stat_function(fun = dchisq, args = list(df = 4)) +
  geom_vline(xintercept = chi_score_mp, color = 'purple')
```

```{r}
(p_value_mp <- 1 - pchisq(chi_score_mp, df = 4))
```

We reject H0 hypothesis in favor of the alternative. There is some correlation
on between pizza toppings and movies.

```{r}
(movies_pizza <- tribble(
  ~Movie, ~Pepperoni, ~Mushrooms, ~Kebab,
  'Frozen 2', 20, 10, 5,
  'Joker', 15, 12, 15,
  'Someone great', 8, 13, 2
))
```
```{r}
movies_pizza %>% select(-Movie) %>%
  chisq.test()
```












