```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Student t-distribution

```{r}
xframe <- seq(-5, 5, length = 100)
df <- 5
dt (xframe, df)

plot(xframe, dt(xframe, df), type = 'l')
```

# One sample t-test

```{r}
kiwi <- read_csv('Downloads/kiwi.csv') %>%
  janitor::clean_names()

head(kiwi, 5)
```

## CLT Conditions

```{r}
kiwi_gs_m <- kiwi %>% 
  filter(species_code == 'GS', gender == 'M')

nrow(kiwi_gs_m)
```
```{r}
ggplot(kiwi_gs_m) +
  geom_histogram(aes(x = height_cm, bin = 30))
```
```{r}
ggplot(kiwi_gs_m) + 
  geom_histogram(aes(x = height_cm, y = ..density..)) +
  stat_function(fun = dnorm, args = list(mean = mean(kiwi_gs_m$height_cm),
                                         sd = sd(kiwi_gs_m$height_cm)))
```

## Setup the hypothesis

H_0: Population mean of male kiwis is 45 cm
H_1: Population mean of male kiwis is not 45 cm

## Assume threshold values

alpha = 0.05

## Calculate the results

```{r}
(point_estimate <- mean(kiwi_gs_m$height_cm))
(sample_sd <- sd(kiwi_gs_m$height_cm))
(SE_gs_m <- sample_sd / sqrt(nrow(kiwi_gs_m)))
(df_gs_m <- nrow(kiwi_gs_m) - 1)
(t_statistics <- (point_estimate - 45) / SE_gs_m)
```
```{r}
ggplot(data.frame(x = seq(-4, 4, length = 200)), aes(x = x)) +
  stat_function(fun = dt, args = list(df = df_gs_m)) +
  geom_segment(aes(x = 0, y = 0, xend = 0, yend = dt(0, df = df_gs_m)),
               color = 'blue', linetype = 'dashed') +
  geom_segment(aes(x = t_statistics, y = 0, xend = t_statistics, 
                   yend = dt(t_statistics, df = df_gs_m)), color = 'red')
```
```{r}
(p_value <- 2 * (1 - pt(t_statistics, df = df_gs_m)))
```

## Draw conclusions

We reject H0 hypothesis in favor of alternative.
Mean Male Great Spotted Kiwi is not 45 cm tall.

```{r}
t.test(kiwi_gs_m$height_cm, mu = 45)
```

# Difference of means t-test

We want to check if mean male Great Spotted are same height as mean male Takoeka.

```{r}
kiwi_gs_t <- kiwi %>% 
  filter(species_code == 'Tok', gender == 'M')
```

## Check CLT conditions

```{r}
nrow(kiwi_gs_t)
nrow(kiwi_gs_m)
```

```{r}
ggplot(kiwi_gs_t) +
  geom_histogram(aes(x = height_cm, y = ..density.., bin = 30)) +
  stat_function(fun = dnorm, args = list(mean = mean(kiwi_gs_t$height_cm), sd = sd(kiwi_gs_t$height_cm)))
```

## Setup the hypothesis

$$H_0: mean_{GS} - mean_{Tokoeka} = 0$$
$$H_1: mean_{GS} - mean_{Tokoeka} \neq 0$$

## Assume threshold values

alpha = 0.05

## Calculate the results

```{r}
(point_estimate <- mean(kiwi_gs_t$height_cm) - mean(kiwi_gs_m$height_cm))
(SE <- sqrt((sd(kiwi_gs_m$height_cm) ^ 2 / nrow(kiwi_gs_m)) + (sd(kiwi_gs_t$height_cm) ^ 2 / nrow(kiwi_gs_t))))
(df <- nrow(kiwi_gs_m) - 1)
(t_statistics <- (point_estimate - 0) / SE)
```
```{r}
ggplot(data.frame(x = seq(-4, 4, length = 100)), aes(x = x)) +
  stat_function(fun = dt, args = list(df = df)) +
  geom_segment(aes(x = t_statistics, y = 0, xend = t_statistics, yend = 0.4,
                   color = 'red'))
```
```{r}
(p_value <- 2 * pt(t_statistics, df = df))
```

## Draw conclusions

We reject H0 hypothesis in favor of alternative.
Mean Male Great Spotted Kiwis is are of different height than Tokoekas.

```{r}
t.test(kiwi_gs_m$height_cm, kiwi_gs_t$height_cm)
```

# Paired data t-test

We want to compare two types of antiperspirant. We had 20 customers try both
formulas and grade them. We want to picke the winner.

```{r}
anti <- read_delim('Downloads/antiper1.dat', delim = ';')
```
```{r}
anti <- anti %>%
  mutate(diff = DryPowderScore - PowderInOilScore)

head(anti, 5)
```

## Check CLT conditions 

- Samples are independent
- Sample size is bigger or equal to 30
```{r}
nrow(anti)
```

- Population distribution is not strongly skewed

```{r}
ggplot(anti) +
  geom_density(aes(x = DryPowderScore), fill = 'brown', alpha = 0.3) +
  geom_density(aes(x = PowderInOilScore), fill = 'black', alpha = 0.3)
```

## Setup the hypothesis

$$H_0: mean_{diff} = 0$$
$$H_1: mean_{diff} \neq 0$$

## Assume threshold values

alpha - significance level - typically 0.05

## Calculate the results

```{r}
(point_estimate <- mean(anti$diff))
(sd_anti <- sd(anti$diff))
(SE_anti <- sd_anti / sqrt(nrow(anti)))
(df_anti <- nrow(anti) - 1)
(t_statistics <- (point_estimate - 0) / SE_anti)
(p_value_anti <- 2 * (1 - pt(t_statistics, df = df_anti)))
```
```{r}
ggplot(data.frame(x = seq(-4, 4, length = 100)), aes(x = x)) +
  stat_function(fun = dt, args = list(df = df_anti)) +
  geom_segment(aes(x = t_statistics, y = 0, xend = t_statistics, yend = 0.4,
                   color = 'red'))
```

## Draw conclusions

We reject H0 hypothesis in favor of alternative. Those two antiperspirants 
don't score the same with customers. Winner is DryPowderScore.
